---
title: "PA_template"
author: "T Huffman"
date: "Sunday, September 14, 2014"
output: html_document
---
# Peer Assignment 1

This assigment is to demonstrate how reprocible process is essential to dat science.

## Loading and processing the data

This first thing to start will be loading of the necessary packages used in the assigment and load the data file used.
```{r,echo=TRUE}
library(reshape)
library(ggplot2)
library(plyr)
options(warn=-1)
activityData<-read.csv("activity.csv", sep=",", stringsAsFactors=F)
```

The dataset has a few zeros and NA's associated with the steps variable. We'll first get an index o the location of the zero entires. 
```{r, echo=TRUE}
zeroIdx<-which(activityData$steps == 0)
head(zeroIdx)
```
Then we'll set them to NA's. create a new data frame and eliminate NA's using na.omit.
```{r, echo=TRUE}
activityData$steps[zeroIdx]<-c(NA)
cleanActivityData<-na.omit(activityData)
```
Now that we have a clean dataset lets create a histogram of the data showing the frequency of step.
```{r, echo=TRUE}
hist(cleanActivityData$steps,breaks=70, xlab="Steps", main="Histogram of Steps")
```

## What is the mean total number of steps taken per day

From this point we'll calculate the mean and median of the steps withn the data frame.
First the mean
```{r, echo=TRUE}
cleanActivityDataMean<-tapply(cleanActivityData$steps,cleanActivityData$date, mean, na.rm=T)
head(cleanActivityDataMean, n=10)
```
Then the median
```{r, echo=TRUE}
cleanActivityDataMedian<-tapply(cleanActivityData$steps,cleanActivityData$date, median, na.rm=T)
head(cleanActivityDataMedian, n=10)
```

## What is the average daily activity pattern

Used tapply to create generate the means associated with the data set
```{r, echo=TRUE}
cleanInterval<-tapply(cleanActivityData$steps,cleanActivityData$interval, mean, na.rm=T)
cleanMeanDF<-data.frame(as.numeric(cleanInterval), dimnames(cleanInterval), stringsAsFactors=F)
names(cleanMeanDF)<-c("steps","interval")
plot(cleanMeanDF$interval,cleanMeanDF$steps, type="l")
```
### Inputing missing values

Using the initial data frame, input mean values where NA's exists.

The folowing reporst the nember of missing values

```{r, echo=TRUE}
sum(!complete.cases(activityData))
```

Now we'll fill in those missing values with the mean value for the observation

```{r, echo=TRUE}
cleanMeans<-ddply(cleanActivityData,.(date,interval), summarise, mean = mean(steps))
for(i in 1:length(activityData$steps))
{
       if(is.na(activityData$steps[i])){
                if(length(cleanMeans$mean[cleanMeans$interval == activityData$interval[i]]) >0){
                        activityData$steps[i]<-cleanMeans$mean[cleanMeans$interval == activityData$interval[i]]
                }else{
                        activityData$steps[i]<-as.numeric(0)
                }        
       }
}
```
Now to recreate a histogram with the missing data added

```{r, echo=TRUE}
newCleanActivity<-activityData[which(activityData$steps !=0),]
newCleanMeans<-tapply(newCleanActivity$steps,newCleanActivity$date, mean, na.rm=T)
newCleanMedian<-tapply(newCleanActivity$steps,newCleanActivity$date, median, na.rm=T)
meansDF<-data.frame(as.numeric(newCleanMeans),dimnames(newCleanMeans))
names(meansDF)<-c("step_means", "date")
hist(meansDF$step_means,breaks=70)
```
## Are there differences in activity patterns between weekdays and weekends?

Finally create a plot that shows the differences between weekdays and weekends

```{r, echo=TRUE}
activityDataFactor<-activityData
activityDataFactor$minutes<-as.factor(activityData$interval)
activityDataFactor$weekdays<-weekdays(as.Date(activityData$date))
activityDataFactor$weekdays<-gsub("Monday|Tuesday|Wednesday|Thursday|Friday","Weekday",activityDataFactor$weekdays)
activityDataFactor$weekdays<-gsub("Saturday|Sunday","Weekend",activityDataFactor$weekdays)

g<-ggplot(activityDataFactor, aes(interval,steps))
g+geom_line()+facet_grid(.~weekdays)
```